---
title: "Cod_DeltaGAM"
author: "Bea Smith"
date: "2/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#1. Load Packages
```{r}
library(tidyverse)
library(mgcv)
library(dplyr)
library(maps)
library(mapdata)
library(sna)
library(phenology)
library(ggplot2)
library(raster)
```

# 2. Explore data, reclass env variables, create Cod dataset
```{r}
total <- read.csv("total.csv")
```

## 2.1 Reclass some ENV Variables

### 2.1.1 Sediment Size
Reclass Sediment to integer
```{r}
unique(total$SEDIMENT)
total$SEDSIZE <- ifelse(total$SEDIMENT ==  '0 - 0.03 Silt/Mud', 3, 
                        ifelse(total$SEDIMENT ==  '0.03 - 0.17 Sand', 4,
                               ifelse(total$SEDIMENT ==  '0.17 - 0.35 Sand', 1, 
                                      ifelse(total$SEDIMENT ==  '0.36 - 0.48 Sand', 5,
                                             ifelse(total$SEDIMENT ==  '0.35 - 0.36 Sand', 6, ifelse(total$SEDIMENT ==  '0.48+ Coarse Sand to Gravel',2,NA))))))

total %>% 
  dplyr::select(SEDIMENT,SEDSIZE) %>% 
  sample_n(5)
```

```{r}
total$SEDIMENT = NULL
total <- total %>%
  dplyr::rename(SEDIMENT = 'SEDSIZE')
```

### 2.1.2 SBFDEPTH
Reclass SBF Depth to integer
```{r}
unique(total$SBFDEPTH) 

total$SBF_RCLS <- ifelse(total$SBFDEPTH ==  'Shallow Depression', 1, 
                        ifelse(total$SBFDEPTH==  'High Flat', 2,
                               ifelse(total$SBFDEPTH ==  'Deeper Flat',3, 
                                      ifelse(total$SBFDEPTH ==  'Steep', 4,
                                             ifelse(total$SBFDEPTH ==  'Low Slope', 5,
                                                    ifelse(total$SBFDEPTH ==  'Somewhat Deep Depression', 6,
                                                           ifelse(total$SBFDEPTH ==  'Moderate Depression', 7,
                                                                  ifelse(total$SBFDEPTH ==  'Shallow Flat', 8,
                                                                         ifelse(total$SBFDEPTH ==  'Somewhat Deep Flat', 9,
                                                                                ifelse(total$SBFDEPTH ==  'Very Shallow Flat',10,
                                                                                       ifelse(total$SBFDEPTH ==  'Deepest Depression',11,
                                                                                              ifelse(total$SBFDEPTH ==  'Deep Depression',12,
                                                                                                     ifelse(total$SBFDEPTH ==  'Moderate Flat',13,
                                                                                                            ifelse(total$SBFDEPTH ==  'Deeper Depression',14,
                                                                                                                   ifelse(total$SBFDEPTH ==  'Side Slope',15,
                                                                                                                          ifelse(total$SBFDEPTH ==  'Very Shallow Depression',16,
                                                                                                                                 ifelse(total$SBFDEPTH ==  'Deepest Flat',17,
                                                                                                                                      ifelse(total$SBFDEPTH ==  'Deep Flat',18,NA))))))))))))))))))

total %>% 
  dplyr::select(SBFDEPTH,SBF_RCLS) %>% 
  sample_n(5)

```
```{r}
total$SBFDEPTH = NULL
total <- total %>%
  dplyr::rename(SBFDEPTH = 'SBF_RCLS')
```

```{r}
total %>% 
  dplyr::select(SBFDEPTH,SEDIMENT) %>% 
  sample_n(5)
```

## 2.2 Selecting ENV variables and subsetting Cod
```{r}
atlantic_cod <- total %>% 
  dplyr::select(GADUS.MORHUA, year, date, lat, lon, depth, SEDIMENT, bottemp, surftemp, surfsalin, botsalin, season, SBFDEPTH) #create new DF with specific environmental variables
```

##2.3 Subsetting by Season
```{r}
Spring_Cod <- subset(atlantic_cod, atlantic_cod$season == "SPRING")

Fall_Cod <- subset(atlantic_cod, atlantic_cod$season == "FALL")
```

##2.4 Splitting into Presence Absence Dataset
```{r}
Spring_Cod_pres <- Spring_Cod
Spring_Cod_pres$GADUS.MORHUA[Spring_Cod_pres$GADUS.MORHUA > 0] <- 1 

Fall_Cod_pres <- Fall_Cod
Fall_Cod_pres$GADUS.MORHUA[Fall_Cod_pres$GADUS.MORHUA > 0] <- 1 

```

Percent of times caught in total trawls
```{r}
Spring_prcnt_caught <- (sum(Spring_Cod_pres$GADUS.MORHUA == 1))/ (nrow(Spring_Cod_pres)) * 100


Fall_prcnt_caught <- (sum(Fall_Cod_pres$GADUS.MORHUA == 1))/ (nrow(Fall_Cod_pres)) * 100
```

##2.5 Caught Fish Datasets 
```{r}
Caught_Spring_Cod <- subset(Spring_Cod, Spring_Cod$GADUS.MORHUA > 0)
  
Caught_Fall_Cod <- subset(Fall_Cod, Fall_Cod$GADUS.MORHUA > 0)
```

#3. Modeling - Spring
##3.1 Pres/Abs 
```{r}
#binomial gam submodel. Binomial answers how many successes we get in a fixed number of trials. Random variable x = # of successes

Spring_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Spring_Cod_pres)

summary(Spring_gam_presabs)
AIC(Spring_gam_presabs)
anova(Spring_gam_presabs)


gam.check(Spring_gam_presabs) #https://www.rdocumentation.org/packages/mgcv/versions/1.2-2/topics/gam.check
```
no bottemp
```{r}
Spring_gam_presabs <- gam(GADUS.MORHUA ~ s(surftemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Spring_Cod_pres)

summary(Spring_gam_presabs)
AIC(Spring_gam_presabs)
anova(Spring_gam_presabs)


gam.check(Spring_gam_presabs)
```
not surf temp
```{r}
Spring_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Spring_Cod_pres)

summary(Spring_gam_presabs)
AIC(Spring_gam_presabs)
anova(Spring_gam_presabs)


gam.check(Spring_gam_presabs)
```
no sediment
```{r}
Spring_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Spring_Cod_pres)

summary(Spring_gam_presabs)
AIC(Spring_gam_presabs)
anova(Spring_gam_presabs)


gam.check(Spring_gam_presabs)
```
no botsalin
```{r}
Spring_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Spring_Cod_pres)

summary(Spring_gam_presabs)
AIC(Spring_gam_presabs)
anova(Spring_gam_presabs)


gam.check(Spring_gam_presabs)
```
no surf salin
```{r}
Spring_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(botsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Spring_Cod_pres)

summary(Spring_gam_presabs)
AIC(Spring_gam_presabs)
anova(Spring_gam_presabs)


gam.check(Spring_gam_presabs)
```
no SBF depth
```{r}
Spring_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin), family = binomial(link = "logit"), data=Spring_Cod_pres)

summary(Spring_gam_presabs)
AIC(Spring_gam_presabs)
anova(Spring_gam_presabs)


gam.check(Spring_gam_presabs)
```

##3.2 Positive CPUE
###3.2.1. Tweedie
```{r}
#p  - the variance of an observation is proportional to its mean to the power p. p must be greater than 1 and less than or equal to 2. 1 would be Poisson, 2 is gamma.
#link - The link function: one of "log", "identity", "inverse", "sqrt", or a power link (Tweedie only). GAM tweedie functions https://stat.ethz.ch/R-manual/R-patched/library/mgcv/html/Tweedie.html

#fit a poisson glm with identity link
Spring_pres.tweedie <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, family = tw(link = "log"), data=Caught_Spring_Cod)

#link = power(0)

summary(Spring_pres.tweedie)
AIC(Spring_pres.tweedie)
anova(Spring_pres.tweedie)

gam.check(Spring_pres.tweedie)
```

###3.2.2 Gamma
```{r}
Spring_pres.gamma <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, family = Gamma(link = "inverse"), data=Caught_Spring_Cod)

summary(Spring_pres.gamma)
AIC(Spring_pres.gamma)
anova(Spring_pres.gamma)

gam.check(Spring_pres.gamma)
```

###3.2.3 Gaussian Log
```{r}
Spring_pres.gamlog <- gam(log(GADUS.MORHUA) ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, data=Caught_Spring_Cod)

summary(Spring_pres.gamlog)
AIC(Spring_pres.gamlog)
anova(Spring_pres.gamlog)


gam.check(Spring_pres.gamlog)
```

###3.2.4 Gaussian Fourth Root 
```{r}
Spring_pres.gam4log <- gam((GADUS.MORHUA)^0.25 ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, data=Caught_Spring_Cod)

summary(Spring_pres.gam4log)
AIC(Spring_pres.gam4log)
anova(Spring_pres.gam4log)


gam.check(Spring_pres.gam4log)
      
```

##3.3 Total Biomass 
###3.3.1 Log Gaussian
```{r}
Spring_total.gamlog <- gam(log(GADUS.MORHUA +1) ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, data=Spring_Cod)

summary(Spring_total.gamlog)
AIC(Spring_total.gamlog)
anova(Spring_total.gamlog)


gam.check(Spring_total.gamlog)
```

###3.3.2 Log Tweedie
```{r}
#p  - the variance of an observation is proportional to its mean to the power p. p must be greater than 1 and less than or equal to 2. 1 would be Poisson, 2 is gamma.
#link - The link function: one of "log", "identity", "inverse", "sqrt", or a power link (Tweedie only). GAM tweedie functions https://stat.ethz.ch/R-manual/R-patched/library/mgcv/html/Tweedie.html

#fit a poisson glm with identity link
Spring_total.tweedie_log <- gam(log(GADUS.MORHUA +1) ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, family = tw, data=Spring_Cod)

#link = power(0)

summary(Spring_total.tweedie_log)
AIC(Spring_total.tweedie_log)
anova(Spring_total.tweedie_log)

gam.check(Spring_total.tweedie_log)
```

###3.3.3 Fourth Root Tweedie  
```{r}
#p  - the variance of an observation is proportional to its mean to the power p. p must be greater than 1 and less than or equal to 2. 1 would be Poisson, 2 is gamma.
#link - The link function: one of "log", "identity", "inverse", "sqrt", or a power link (Tweedie only). GAM tweedie functions https://stat.ethz.ch/R-manual/R-patched/library/mgcv/html/Tweedie.html

#fit a poisson glm with identity link
Spring_total.tweedie_fr <- gam((GADUS.MORHUA)^.25 ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = tw, data=Spring_Cod)

#link = power(0)

summary(Spring_total.tweedie_fr)
AIC(Spring_total.tweedie_fr)
anova(Spring_total.tweedie_fr)

gam.check(Spring_total.tweedie_fr)
```
### 3.3.4 Fourth Root Gaussian 
```{r}
Spring_total.gamlog <- gam((GADUS.MORHUA)^.25 ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), data=Spring_Cod)

summary(Spring_total.gamlog)
AIC(Spring_total.gamlog)
anova(Spring_total.gamlog)

#window(width=8, height=8, record=T) #code to see how the model fits
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Spring_total.gamlog)
```

##Covariate Plot

```{r}
#dev.new(width=9,height=9,record=T)
par(mfrow=c(2,3),mai=c(0.4,0.5,0.1,0.1),omi=c(0.1,0,0.01,0.01))

#SurfSalin
grid.res <- 500 # estimation grid resolution
grid.bin <- data.frame(
  "bottemp"=mean(Spring_Cod$bottemp, na.rm = TRUE),
  "surftemp" = mean(Spring_Cod$surftemp, na.rm = TRUE),
  "surfsalin"=seq(min(Spring_Cod$surfsalin, na.rm = TRUE), max(Spring_Cod$surfsalin, na.rm = TRUE),length=grid.res),
  "botsalin"=mean(Spring_Cod$botsalin, na.rm = TRUE),
  "SEDIMENT"= (Spring_Cod$SEDIMENT = "2" ),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"))

pred.bin <- predict(Spring_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res,ncol=niter)
for(i in 1:grid.res){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean, na.rm = TRUE)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975), na.rm = TRUE)
plot(grid.bin$surfsalin, pred.mean, type="l", ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$surfsalin, pred.95CI[1,], lty=2)
lines(grid.bin$surfsalin, pred.95CI[2,], lty=2)
legend("topleft",legend="a. Surface Salinity",cex=1.4,bty="n")

#Botsalin

grid.bin <- data.frame(
  "bottemp"=mean(Spring_Cod$bottemp, na.rm=TRUE),
  "surftemp" = mean(Spring_Cod$surftemp, na.rm = TRUE),
  "surfsalin"=mean(Spring_Cod$surfsalin, na.rm=TRUE),
  "botsalin"=seq(min(Spring_Cod$botsalin, na.rm=TRUE), max(Spring_Cod$botsalin, na.rm=TRUE),length=grid.res),
  "SEDIMENT"= (Spring_Cod$SEDIMENT = "2"),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"))
  

pred.bin <- predict(Spring_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res,ncol=niter)
for(i in 1:grid.res){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975))
plot(grid.bin$botsalin, pred.mean, type="l", ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$botsalin, pred.95CI[1,], lty=2)
lines(grid.bin$botsalin, pred.95CI[2,], lty=2)
legend("topleft",legend="b. Bottom Salinity",cex=1.4,bty="n")

#Surftemp

grid.bin <- data.frame(
  "bottemp"=mean(Spring_Cod$bottemp, na.rm=TRUE),
  "surftemp" =seq(min(Spring_Cod$surftemp, na.rm=TRUE), max(Spring_Cod$surftemp, na.rm=TRUE),length=grid.res),
  "surfsalin"=mean(Spring_Cod$surfsalin, na.rm = TRUE),
  "botsalin"=mean(Spring_Cod$botsalin, na.rm=TRUE),
  "SEDIMENT"= (Spring_Cod$SEDIMENT = "2" ),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"))
  

pred.bin <- predict(Spring_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res,ncol=niter)
for(i in 1:grid.res){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975))
plot(grid.bin$surftemp, pred.mean, type="l", ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$surftemp, pred.95CI[1,], lty=2)
lines(grid.bin$surftemp, pred.95CI[2,], lty=2)
legend("topleft",legend="c. Surface Temp",cex=1.4,bty="n")


#Bottemp

grid.bin <- data.frame(
  "bottemp"=seq(min(Spring_Cod$bottemp, na.rm=TRUE), max(Spring_Cod$bottemp, na.rm=TRUE),length=grid.res),
  "surftemp" = mean(Spring_Cod$surftemp, na.rm = TRUE),
  "surfsalin"=mean(Spring_Cod$surfsalin, na.rm = TRUE),
  "botsalin"=mean(Spring_Cod$botsalin, na.rm=TRUE),
  "SEDIMENT"= (Spring_Cod$SEDIMENT = "2" ),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"))
  

pred.bin <- predict(Spring_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res,ncol=niter)
for(i in 1:grid.res){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975))
plot(grid.bin$bottemp, pred.mean, type="l", ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$bottemp, pred.95CI[1,], lty=2)
lines(grid.bin$bottemp, pred.95CI[2,], lty=2)
legend("topleft",legend="d. Bottom Temp",cex=1.4,bty="n")
mtext("Probability of the Presence of Cod",side=2,outer=F, adj = 0, line=2.3,cex=1)

#SEDIMENT
grid.res.sed <- 6
grid.bin <- data.frame(
  "bottemp"=mean(Spring_Cod$bottemp, na.rm=TRUE),
  "surftemp" = mean(Spring_Cod$surftemp, na.rm = TRUE),
  "surfsalin"=mean(Spring_Cod$surfsalin, na.rm=TRUE),
  "botsalin"=mean(Spring_Cod$botsalin, na.rm=TRUE),
  "SEDIMENT"= c("3", "4", "1", "6", "5", "2"),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"))
 

pred.bin <- predict(Spring_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res.sed,ncol=niter)
for(i in 1:grid.res.sed){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean, na.rm = TRUE)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975), na.rm = TRUE)
grid.bin$SEDIMENT <- factor(grid.bin$SEDIMENT ,levels=c("3", "4", "1", "6", "5", "2"),labels=c("0.03", "0.17", "0.35", "0.36", "0.48", "0.5"))
plot(grid.bin$SEDIMENT, pred.mean, type="p", mgp=c(2,0.7,0), ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",boxwex=0.01,cex.axis=1.1,xaxt="n")
points(grid.bin$SEDIMENT,pred.mean, type="p",cex=1.5,pch=20)
axis(1,at=c(1,2,3,4,5,6),labels=c("0.03", "0.17", "0.35", "0.36", "0.48", "0.5"),cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$SEDIMENT, pred.95CI[1,], lty=2)
lines(grid.bin$SEDIMENT, pred.95CI[2,], lty=2)
legend("topleft",legend="e. Sediment Size",cex=1.4,bty="n")

#SBFDEPTH
grid.res.SBF <- 18
grid.bin <- data.frame(
  "bottemp"=mean(Spring_Cod$bottemp, na.rm=TRUE),
  "surftemp" = mean(Spring_Cod$surftemp, na.rm = TRUE),
  "surfsalin"=mean(Spring_Cod$surfsalin, na.rm=TRUE),
  "botsalin"=mean(Spring_Cod$botsalin, na.rm=TRUE),
  "SEDIMENT"= (Spring_Cod$SEDIMENT = "2"),
  "SBFDEPTH" = c("1", "7", "2","13", "9", "14", "3", "5", "8", "15", "4", "17", "6", "18", "11", "10", "16", "12"))
  
pred.bin <- predict(Spring_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res.SBF,ncol=niter)
for(i in 1:grid.res.SBF){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean, na.rm = TRUE)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975), na.rm = TRUE)
grid.bin$SBFDEPTH <- factor(grid.bin$SBFDEPTH ,levels=c("1", "7", "2","13", "9", "14", "3", "5", "8", "15", "4", "17", "6", "18", "11", "10", "16", "12"))
plot(grid.bin$SBFDEPTH, pred.mean, type="p", mgp=c(2,0.7,0), xlim=c(1,18), ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",boxwex=0.01,cex.axis=1.1,xaxt="n")
points(grid.bin$SBFDEPTH,pred.mean, type="p",cex=1.5,pch=20)
axis(1,at=c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18),labels=c("Shllw Dep.", "Mdrte Dep", "High Flat","Moderate Flat", "Smwht Deep Flat", "Deeper Dep", "Deeper Flat", "Low Slope", "Shallow Flat", "Side Slope", "Steep", "Deepest Flat", "Smwht Deep Dep", "Deep Flat", "Deepest Dep", "Vry Shallow Flat", "Vry Shallow Dep", "Deep Dep"), cex.axis=.4,mgp=c(2,0.7,0), las = 2)
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$SBFDEPTH, pred.95CI[1,], lty=2)
lines(grid.bin$SBFDEPTH, pred.95CI[2,], lty=2)
legend("topleft",legend="f. Benthic Habitat",cex=1.4,bty="n")

```

```{r}
dev.new(width=12,height=8,record=T)
par(mfrow=c(2,3),mai=c(0.3,0.3,0,0),omi=c(0.35,0.35,0.1,0.1))
plot(Spring_total.tweedie_fr, shade=TRUE, seWithMean=TRUE, scale=0, all.terms=TRUE)
```


#4. Modeling - Fall
##4.1 Pres/Abs 
```{r}
#binomial gam submodel. Binomial answers how many successes we get in a fixed number of trials. Random variable x = # of successes

Fall_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Fall_Cod_pres)

summary(Fall_gam_presabs)
AIC(Fall_gam_presabs)
anova(Fall_gam_presabs)

#code to see how the model fits 
#windows(width=8,height=8,record=T) #code to see how the model fits 
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Fall_gam_presabs)
```
no bot temp
```{r}
#binomial gam submodel. Binomial answers how many successes we get in a fixed number of trials. Random variable x = # of successes

Fall_gam_presabs <- gam(GADUS.MORHUA ~ s(surftemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Fall_Cod_pres)

summary(Fall_gam_presabs)
AIC(Fall_gam_presabs)
anova(Fall_gam_presabs)

#code to see how the model fits 
#windows(width=8,height=8,record=T) #code to see how the model fits 
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Fall_gam_presabs)
```
no surf temp
```{r}
#binomial gam submodel. Binomial answers how many successes we get in a fixed number of trials. Random variable x = # of successes

Fall_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Fall_Cod_pres)

summary(Fall_gam_presabs)
AIC(Fall_gam_presabs)
anova(Fall_gam_presabs)

#code to see how the model fits 
#windows(width=8,height=8,record=T) #code to see how the model fits 
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Fall_gam_presabs)
```
no sediment
```{r}
#binomial gam submodel. Binomial answers how many successes we get in a fixed number of trials. Random variable x = # of successes

Fall_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Fall_Cod_pres)

summary(Fall_gam_presabs)
AIC(Fall_gam_presabs)
anova(Fall_gam_presabs)

#code to see how the model fits 
#windows(width=8,height=8,record=T) #code to see how the model fits 
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Fall_gam_presabs)
```
no bot salin
```{r}
#binomial gam submodel. Binomial answers how many successes we get in a fixed number of trials. Random variable x = # of successes

Fall_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(surfsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Fall_Cod_pres)

summary(Fall_gam_presabs)
AIC(Fall_gam_presabs)
anova(Fall_gam_presabs)

#code to see how the model fits 
#windows(width=8,height=8,record=T) #code to see how the model fits 
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Fall_gam_presabs)
```
no surf salin
```{r}
#binomial gam submodel. Binomial answers how many successes we get in a fixed number of trials. Random variable x = # of successes

Fall_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(botsalin) + factor(SBFDEPTH), family = binomial(link = "logit"), data=Fall_Cod_pres)

summary(Fall_gam_presabs)
AIC(Fall_gam_presabs)
anova(Fall_gam_presabs)

#code to see how the model fits 
#windows(width=8,height=8,record=T) #code to see how the model fits 
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Fall_gam_presabs)
```
no SBF depth
```{r}
#binomial gam submodel. Binomial answers how many successes we get in a fixed number of trials. Random variable x = # of successes

Fall_gam_presabs <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin), family = binomial(link = "logit"), data=Fall_Cod_pres)

summary(Fall_gam_presabs)
AIC(Fall_gam_presabs)
anova(Fall_gam_presabs)

#code to see how the model fits 
#windows(width=8,height=8,record=T) #code to see how the model fits 
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Fall_gam_presabs)
```

##4.2 Positive CPUE
###4.2.1. Tweedie
```{r}
#p  - the variance of an observation is proportional to its mean to the power p. p must be greater than 1 and less than or equal to 2. 1 would be Poisson, 2 is gamma.
#link - The link function: one of "log", "identity", "inverse", "sqrt", or a power link (Tweedie only). GAM tweedie functions https://stat.ethz.ch/R-manual/R-patched/library/mgcv/html/Tweedie.html

#fit a poisson glm with identity link
Fall_pres.tweedie <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, family = tw(link = "log"), data=Caught_Fall_Cod)

#link = power(0)

summary(Fall_pres.tweedie)
AIC(Fall_pres.tweedie)
anova(Fall_pres.tweedie)

gam.check(Fall_pres.tweedie)
```
###4.2.2 Gamma
```{r}
Fall_pres.gamma <- gam(GADUS.MORHUA ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, family = Gamma(link = "inverse"), data=Caught_Fall_Cod)

summary(Fall_pres.gamma)
AIC(Fall_pres.gamma)
anova(Fall_pres.gamma)

gam.check(Fall_pres.gamma)
```
###4.2.3 Gaussian Log
```{r}
Fall_pres.gamlog <- gam(log(GADUS.MORHUA) ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, data=Caught_Fall_Cod)

summary(Fall_pres.gamlog)
AIC(Fall_pres.gamlog)
anova(Fall_pres.gamlog)

#window(width=8, height=8, record=T) #code to see how the model fits
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Fall_pres.gamlog)
```
###4.2.4 Gaussian Fourth Root 
```{r}
Fall_pres.gam4log <- gam((log(GADUS.MORHUA)^0.25) ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, data=Caught_Fall_Cod)


summary(Fall_pres.gam4log)
AIC(Fall_pres.gam4log)
anova(Fall_pres.gam4log)

#window(width=8, height=8, record=T) #code to see how the model fits
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))

#https://www.rdocumentation.org/packages/mgcv/versions/1.2-2/topics/gam.check
gam.check(Fall_pres.gam4log)
      
```

##4.3 Total Biomass 
### 4.3.1 Fourth Root Tweedie

```{r}
#p  - the variance of an observation is proportional to its mean to the power p. p must be greater than 1 and less than or equal to 2. 1 would be Poisson, 2 is gamma.
#link - The link function: one of "log", "identity", "inverse", "sqrt", or a power link (Tweedie only). GAM tweedie functions https://stat.ethz.ch/R-manual/R-patched/library/mgcv/html/Tweedie.html

#fit a poisson glm with identity link
Fall_total.tweedie_fr <- gam((GADUS.MORHUA)^.25 ~ s(bottemp) + s(surftemp) + factor(SEDIMENT) + s(botsalin) + s(surfsalin) + factor(SBFDEPTH), family = tw, data=Fall_Cod)

#link = power(0)

summary(Fall_total.tweedie_fr)
AIC(Fall_total.tweedie_fr)
anova(Fall_total.tweedie_fr)

gam.check(Fall_total.tweedie_fr)
```
### 4.3.2 Fourth Root Gaussian 
```{r}
Fall_total.gamlog <- gam((GADUS.MORHUA)^.25 ~ s(bottemp) + s(surftemp) + SEDIMENT + s(botsalin) + s(surfsalin) + SBFDEPTH, data=Fall_Cod)

summary(Fall_total.gamlog)
AIC(Fall_total.gamlog)
anova(Fall_total.gamlog)

#window(width=8, height=8, record=T) #code to see how the model fits
#par(mfrow=c(2,2),mai=c(0.3,0.3,0.3,0.3),omi=c(0.1,0.1,0.1,0.1))
gam.check(Fall_total.gamlog)
```

##Covariate Plot

```{r}
#dev.new(width=9,height=9,record=T)
par(mfrow=c(2,3),mai=c(0.4,0.5,0.1,0.1),omi=c(0.1,0,0.01,0.01))

#Surftemp
grid.res <- 500 # estimation grid resolution
grid.bin <- data.frame(
  "bottemp"=mean(Fall_Cod$bottemp, na.rm = TRUE),
  "surftemp"=seq(min(Fall_Cod$surftemp, na.rm = TRUE), max(Fall_Cod$surftemp, na.rm = TRUE),length=grid.res),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"),
  "surfsalin"=mean(Fall_Cod$surfsalin, na.rm = TRUE),
  "botsalin"=mean(Fall_Cod$botsalin, na.rm = TRUE),
  "SEDIMENT"= (Fall_Cod$SEDIMENT = "2"))
  

pred.bin <- predict(Fall_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res,ncol=niter)
for(i in 1:grid.res){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean, na.rm = TRUE)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975), na.rm = TRUE)
plot(grid.bin$surftemp, pred.mean, type="l", ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$surftemp, pred.95CI[1,], lty=2)
lines(grid.bin$surftemp, pred.95CI[2,], lty=2)
legend("topleft",legend="a. Surface Temp",cex=1.4,bty="n")

#Bottemp

grid.bin <- data.frame(
  "bottemp"=seq(min(Fall_Cod$bottemp, na.rm=TRUE), max(Fall_Cod$bottemp, na.rm=TRUE),length=grid.res),
  "surftemp"=mean(Fall_Cod$surftemp, na.rm = TRUE),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"),
  "surfsalin"=mean(Fall_Cod$surfsalin, na.rm=TRUE),
  "botsalin"=mean(Fall_Cod$botsalin, na.rm=TRUE),
  "SEDIMENT"= (Fall_Cod$SEDIMENT = "2"))
  

pred.bin <- predict(Fall_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res,ncol=niter)
for(i in 1:grid.res){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975))
plot(grid.bin$bottemp, pred.mean, type="l", ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$bottemp, pred.95CI[1,], lty=2)
lines(grid.bin$bottemp, pred.95CI[2,], lty=2)
legend("topleft",legend="b. Bottom Temp",cex=1.4,bty="n")

#Surfsalin

grid.bin <- data.frame(
  "bottemp"=mean(Fall_Cod$bottemp, na.rm=TRUE),
  "surftemp"=mean(Fall_Cod$surftemp, na.rm=TRUE),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"),
  "surfsalin"=seq(min(Fall_Cod$surfsalin, na.rm=TRUE), max(Fall_Cod$surfsalin, na.rm=TRUE),length=grid.res),
  "botsalin"=mean(Fall_Cod$botsalin, na.rm=TRUE),
  "SEDIMENT"= (Fall_Cod$SEDIMENT = "2"))
  

pred.bin <- predict(Fall_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res,ncol=niter)
for(i in 1:grid.res){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975))
plot(grid.bin$surfsalin, pred.mean, type="l", ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$surfsalin, pred.95CI[1,], lty=2)
lines(grid.bin$surfsalin, pred.95CI[2,], lty=2)
legend("topleft",legend="c. Surface Salinity",cex=1.4,bty="n")

#Botsalin

grid.bin <- data.frame(
  "bottemp"=mean(Fall_Cod$bottemp, na.rm=TRUE),
  "surftemp"=mean(Fall_Cod$surftemp, na.rm=TRUE),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"),
  "surfsalin"=mean(Fall_Cod$surfsalin, na.rm=TRUE),
  "botsalin"=seq(min(Fall_Cod$botsalin, na.rm=TRUE), max(Fall_Cod$botsalin, na.rm=TRUE),length=grid.res),
  "SEDIMENT"= (Fall_Cod$SEDIMENT = "2"))
  

pred.bin <- predict(Fall_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res,ncol=niter)
for(i in 1:grid.res){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975))
plot(grid.bin$botsalin, pred.mean, type="l", ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$botsalin, pred.95CI[1,], lty=2)
lines(grid.bin$botsalin, pred.95CI[2,], lty=2)
legend("topleft",legend="d. Bottom Salinity",cex=1.4,bty="n")
mtext("Probability of the Presence of Cod",side=2,outer=F, adj = 0, line=2.3,cex=1)

#SEDIMENT
grid.res.sed <- 6
grid.bin <- data.frame(
  "bottemp"=mean(Fall_Cod$bottemp, na.rm=TRUE),
  "surftemp"=mean(Fall_Cod$surftemp, na.rm=TRUE),
  "SBFDEPTH" = (Spring_Cod$SBFDEPTH = "13"),
  "surfsalin"=mean(Fall_Cod$surfsalin, na.rm=TRUE),
  "botsalin"=mean(Fall_Cod$botsalin, na.rm=TRUE),
  "SEDIMENT"= c("3", "4", "1", "6", "5", "2"))
 

pred.bin <- predict(Fall_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res.sed,ncol=niter)
for(i in 1:grid.res.sed){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean, na.rm = TRUE)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975), na.rm = TRUE)
grid.bin$SEDIMENT <- factor(grid.bin$SEDIMENT ,levels=c("3", "4", "1", "6", "5", "2"),labels=c("0.03", "0.17", "0.35", "0.36", "0.48", "0.5"))
plot(grid.bin$SEDIMENT, pred.mean, type="p", mgp=c(2,0.7,0), ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",boxwex=0.01,cex.axis=1.1,xaxt="n")
points(grid.bin$SEDIMENT,pred.mean, type="p",cex=1.5,pch=19)
axis(1,at=c(1,2,3,4,5,6),labels=c("0.03", "0.17", "0.35", "0.36", "0.48", "0.5"),cex.axis=1.1,mgp=c(2,0.7,0))
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$SEDIMENT, pred.95CI[1,], lty=2)
lines(grid.bin$SEDIMENT, pred.95CI[2,], lty=2)
legend("topleft",legend="e. Sediment Size",cex=1.4,bty="n")


#SBFDEPTH
grid.res.SBF <- 18
grid.bin <- data.frame(
  "bottemp"=mean(Fall_Cod$bottemp, na.rm=TRUE),
  "surftemp"=mean(Fall_Cod$surftemp, na.rm=TRUE),
  "surfsalin"=mean(Fall_Cod$surfsalin, na.rm=TRUE),
  "botsalin"=mean(Fall_Cod$botsalin, na.rm=TRUE),
  "SEDIMENT"= (Fall_Cod$SEDIMENT = "2"),
  "SBFDEPTH" = c("1", "7", "2","13", "9", "14", "3", "5", "8", "15", "4", "17", "6", "18", "11", "10", "16", "12"))
  
pred.bin <- predict(Fall_gam_presabs, grid.bin, type="response", se=T)
niter <- 10000  # number of replicates
pred.bin.boot <- matrix(NA,nrow=grid.res.SBF,ncol=niter)
for(i in 1:grid.res.SBF){
  pred.bin.boot[i,] <- rnorm(niter, pred.bin$fit[i], pred.bin$se.fit[i])}
pred.mean <- apply(pred.bin.boot,1,mean, na.rm = TRUE)
pred.95CI <- apply(pred.bin.boot,1,quantile,probs=c(0.025,0.975), na.rm = TRUE)
grid.bin$SBFDEPTH <- factor(grid.bin$SBFDEPTH ,levels=c("1", "7", "2","13", "9", "14", "3", "5", "8", "15", "4", "17", "6", "18", "11", "10", "16", "12"))
plot(grid.bin$SBFDEPTH, pred.mean, type="p", mgp=c(2,0.7,0), xlim=c(1,18), ylim=c(0,1),xlab="",lwd=2,cex.lab=1.4,ylab="",boxwex=0.01,cex.axis=1.1,xaxt="n")
points(grid.bin$SBFDEPTH,pred.mean, type="p",cex=1.5,pch=20)
axis(1,at=c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18),labels=c("Shllw Dep.", "Mdrte Dep", "High Flat","Moderate Flat", "Smwht Deep Flat", "Deeper Dep", "Deeper Flat", "Low Slope", "Shllw Flat", "Side Slope", "Steep", "Deepest Flat", "Smwht Deep Dep", "Deep Flat", "Deepest Dep", "Vry Shllw Flat", "Vry Shllw Dep", "Deep Dep"), cex.axis=.4,mgp=c(2,0.7,0), las = 2)
#axis(2,at=c(0,0.40,0.80),labels=c("0.0","0.4","0.8"),cex.axis=1.1,mgp=c(2,0.7,0))
lines(grid.bin$SBFDEPTH, pred.95CI[1,], lty=2)
lines(grid.bin$SBFDEPTH, pred.95CI[2,], lty=2)
legend("topleft",legend="f. Benthic Habitat",cex=1.4,bty="n")


```

```{r}
dev.new(width=12,height=8,record=T)
par(mfrow=c(2,3),mai=c(0.3,0.3,0,0),omi=c(0.35,0.35,0.1,0.1))
plot(Fall_total.tweedie_fr, shade=TRUE, seWithMean=TRUE, scale=0, all.terms=TRUE)
```

#5. Building habitat predictions from models 
##5.1 Raster Paths
```{r}
## Raster paths
surfsalin <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Spring/surfsalin_final.tif")
surfsalin
surftemp <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Spring/surftemp_final.tif")
surftemp
bottemp <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Spring/bottemp_final.tif")
bottemp
botsalin <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Spring/botsalin_final.tif")
botsalin

SEDIMENT <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Spring/SEDfloat_final.tif")
SEDIMENT
SBFDEPTH <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Spring/SBFfloat_final.tif")
SBFDEPTH


```

##5.2 Spring Prediction
Raster Stack Spring
```{r}
Spring_rs_stack <- stack(bottemp, surftemp, SEDIMENT, botsalin, surfsalin, SBFDEPTH)
names(Spring_rs_stack) <- c("bottemp", "surftemp", "SEDIMENT", "botsalin", "surfsalin","SBFDEPTH")


raster::stack(Spring_rs_stack)
plot(Spring_rs_stack, col = RColorBrewer::brewer.pal(11, "RdBu"))

```

Log odds/logit surface
```{r}
loglogit_raster = predict(Spring_rs_stack,Spring_gam_presabs,type = "link")
plot(loglogit_raster, col = RColorBrewer::brewer.pal(11, "RdBu"))
writeRaster(loglogit_raster,filename = "loglogitspring_gam.tif",overwrite = TRUE)
```

Habitat Prob. Surface
```{r}
Spring_prob_raster = predict(Spring_rs_stack, Spring_gam_presabs, type="response")
plot(Spring_prob_raster, col = RColorBrewer::brewer.pal(11, "RdBu"))
writeRaster(Spring_prob_raster,filename = "probspring_gam.tif",overwrite = TRUE)
```

Binary Classification (Prob > .5)
```{r}
Spring_habitat_raster <- reclassify(Spring_prob_raster, c(-Inf,0.5,0, 0.5,1,1))
plot(Spring_habitat_raster, col = RColorBrewer::brewer.pal(11, "RdBu"))
writeRaster(Spring_habitat_raster,filename = "habitatspring_gam.tif",overwrite = TRUE)
```


##5.3 Fall Prediction 
```{r}
surfsalin_fall <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Fall/surfsalin_fall.tif")
surftemp_fall <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Fall/surftemp_fall.tif")
bottemp_fall <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Fall/bottemp_fall.tif")
botsalin_fall <- raster("C:/Users/bs315/Desktop/ENV Rasters/MyProject/Fall/botsalin_fall.tif")
```

Raster Stack Fall
```{r}
# Change to reading in the raster stack
Fall_rs_stack <- stack(bottemp_fall, surftemp_fall, SEDIMENT, botsalin_fall, surfsalin_fall,SBFDEPTH)
names(Fall_rs_stack) <- c("bottemp", "surftemp", "SEDIMENT", "botsalin", "surfsalin", "SBFDEPTH")


raster::stack(Fall_rs_stack)
plot(Fall_rs_stack, col = RColorBrewer::brewer.pal(11, "RdBu"))
```

Log odds/logit surface
```{r}
Fall_loglogit_raster = predict(Fall_rs_stack, Fall_gam_presabs, type="link")
plot(Fall_loglogit_raster, col = RColorBrewer::brewer.pal(11, "RdBu"))
writeRaster(Fall_loglogit_raster,filename = "loglogitfall_gam.tif",overwrite = TRUE)

```

Habitat Prob. Surface
```{r}
Fall_prob_raster = predict(Fall_rs_stack, Fall_gam_presabs, type="response")
plot(Fall_prob_raster, col = RColorBrewer::brewer.pal(11, "RdBu"))
writeRaster(Fall_prob_raster,filename = "probfall_gam.tif", overwrite = TRUE)
```

Binary Classification (Prob > .5)
```{r}
Fall_habitat_raster <- reclassify(Fall_prob_raster, c(-Inf,0.5,0, 0.5,1,1))
plot(Fall_habitat_raster, col = RColorBrewer::brewer.pal(11, "RdBu"))
writeRaster(Fall_habitat_raster,filename = "habitatfall_gam.tif", overwrite = TRUE)
```
